---
title: "Filters and Joins"
author: "Ben Parkhurst"
date: "January 28, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
library(kableExtra)
library(readxl)
library(here)
```
Get some fish.csv and kelp_fronds.xlsx data
Note all of the sites are located into different tabs
It defaults to read teh first tab but you can add an argument in sheet = 

```{r}
fish <- read_csv(here("data", "fish.csv"))
kelp_abur <- read_excel(here("data", "kelp_fronds.xlsx"),
                        sheet = "abur")
```
###Filter by an exact match

Exact match: `==`
Note that you could also filter using <><= etc

Example 1: only keep observatinos from `fish` where the common_name variable name matches (==) "garibaldi"

```{r}
fish_garibaldi <- fish %>% 
  filter(common_name == "garibaldi")
```
Activity: create a subset called fish_over50 starting from fish, total_count >=50

```{r}
fish_over50 <- fish %>% 
  filter(total_count >= 50)
```
### What if we want to filter by this OR this?

You can use the vertical line `|`for or

```{r}
fish_2sp <- fish %>% 
  filter(common_name == "garibaldi" | common_name == "blacksmith")
```
#If you're filtering for multiple matches within the same variable, you can use %in%

```{r}
fish_3sp <- fish %>% 
  filter(common_name %in% c("garibaldi","blacksmith", "black surfperch"))
```
Activity: create a subset from fish that creates all observations all observations frmo 2016 or the sommon name is garibaldi:

```{r}
fish_gar_2016 <- fish %>% 
  filter(common_name == "garibaldi" | year == 2016)
```
filter to match this and that:
can use a `,` or `&`

```{r}
aque_2018 <- fish %>%
  filter(year == 2018, site =="aque")
```
To exclude things: `!=`

### stringr::str_detect()

```{r}
fish_black <- fish %>% 
  filter(str_detect(common_name, pattern= "black"))

```
Subset of data if the common name must include the string "it"

```{r}
fish_it <- fish %>% 
  filter(str_detect(common_name, pattern= "it"))
```
### `dplyr:*_join`

```{r}
ab_kelpfish <- kelp_abur %>% 
  full_join(fish, by = c("year","site"))

```
### left_join() to keep all things from x, and merge the matchings things from y with it (exclude things from y that don't have a match in x)

```{r}
kelp_fish_left <- kelp_abur %>% 
  left_join(fish, by = c("year", "site"))
```
### Inner_join to merge data frames.  ONly will keep the observations that match both

```{r}
kelp_fish_inner <- kelp_abur %>% 
  inner_join(fish, by = c("year", "site"))
```
### Pipe together multiple steps


- create an object names `my_fish_join`
- starting from `fish`
- filtering to only including observations at site "abur" from the year 2017
- join the kelp_abur data frame to the resulting subset using left_join()
-  add a new column using mutate() to calcuate fih per kelp frond density counts of fish / total fronds

```{r}
my_fish_join <- fish %>%
  filter(year == 2017, site =="abur") %>% 
  left_join(kelp_abur, by = c("year", "site")) %>% 
  mutate(fish_per_frond = total_count/total_fronds)


```

```{r}
my_fish_join %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped", 
                full_width = FALSE)
  
```
Make a graph
```{r}
ggplot(data = my_fish_join, aes(x = common_name, y = fish_per_frond)) + 
  geom_col(aes(fill = common_name))
```

