---
title: "Pivot Lobsters"
author: "Ben Parkhurst"
date: "January 27, 2020"
output: html_document
---


Install Packages and read in lobster data:
```{r}
library(tidyverse)
library(readxl)
library(here)
library(skimr) #install.packages("skimr")
library(kableExtra) #install.packages("kableExtra")

#read in the lobster data
#lobsters <- read_xlsx(here("data", "lobsters.xlsx"),
 #                     skip = 4)

#Sent wrong data
#read in the lobster data
lobsters <- read_xlsx(here("data", "lobsters2.xlsx"),
                      skip = 4)

```
Viewing our data with skimmer
```{r}
skimr::skim(lobsters)
#skim(lobsters would be equivalent code)

```
#Pivot Tables
In R, Pivot tables are: `group_by() %>% summarize()`
`n()` function will count observations

```{r}
#group by year
lobsters %>% 
  group_by(year) %>% 
  summarize(count_by_year = n())
#group by year and site
lobsters %>% 
  group_by(year,site) %>% 
  summarize(count_by_year_site = n())
#group by site and year
lobsters %>% 
  group_by(site,year) %>% 
  summarize(count_by_year_site = n())
```
Summarize by multiple variable names
And assign as an object
```{r}
siteyear_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE))
```
Table formatting with `kableExtra::Kable()`
```{r}
siteyear_summary %>% 
  kable()
```

There are now `r nrow(lobsters)` total lobsters included in this report

## Activity
1 build from our analyses: calculate the median lobster size for each site and year (hint: `size_mm` and a function called `median`
2. create a ggplot and save. For example:
3. sync this to github

```{r}
siteyear_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_site_year = n(),
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE),
            median_size_mm = median(size_mm, na.rm = TRUE))

ggplot(data = siteyear_summary, aes(x=year, y=median_size_mm, color = site)) + 
  geom_line()

```
Table formatting with `kableExtra::Kable()`
```{r}
siteyear_summary %>% 
  kable()
```

