---
title: "Creating Graphs with ggplot2"
author: "Ben Parkhurst"
date: "January 27, 2020"
output: html_document
---
# Plots with ggplot2

We are going to make plots in R using ggplot2 package from tidyverse.

We will install these packages:

- tidyverse - suite of packages for wangling and visualization (ggplot)
- readxl - to read in excel files (in many ways)
- here - help with files paths

```{r setup}
library(tidyverse) #install.packages("tidyverse")
library(here)##install.packages("here")
library(readxl) #install.packages ("readxl")

```
# California National Parks Data

```{r read data}

ca_np <- read_csv(here("data/ca_np.csv"))

ci_np <- read_excel(here("data", "ci_np.xlsx"))
```
# nake my own first graph
We're going to a make a line plot of year (x-variable) vs visitors (y-variable) in ggplot:

```{r plot}

ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()

#can use geom_point or goem_area for different types of graphs

```
Make a base layer for my CI plots:
```{r}
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()

gg_base + geom_point()
```
# Now let's start customizing!
- line / point colors: color
- change line type: linetype
-change point style: pch
- Change transparency: alpha
```{r}
gg_base + 
  geom_line(
    color = "purple"
  )

```
Activity: create a ggplot scatterplot (geom_point) that has a triangle symbols (hint: ?pch), of a non-default size (size=#) and that are updated color.

```{r}
gg_base + 
  geom_point(
    color = "red",
    pch = 2,
    size = 4
  
  )


```
# Update graph aesthetics based on variables

```{r}
gg_base +
  geom_point(
    aes(size = visitors,
    color = year)
  )
#add an argument to update the color based on the year value

```
# Use gg-theme for overall updates
- theme_light()
- theme_minimal()
- theme_bw()
```{r}
gg_base + 
  geom_area() +
  theme_linedraw()

```
# Create a multi-series line plot using data for all CA national parks
```{r}
ggplot(data = ca_np, aes(x=year, y = visitors, group = park_name, color = park_name)) +
  geom_line() +
  theme_get()

```
#this code will also work:
```{r}
my_np_graph <- ggplot(data = ca_np, aes(x = year, y= visitors)) + 
  geom_point(aes(color = park_name), show.legend = FALSE) + 
  facet_wrap(~park_name, scales = "free") +
  theme_minimal()

```
Export my graph as a .png
```{r}
ggsave(plot = my_np_graph, here("figures", "np_graph.png"))
```

