---
title: "Tidying"
author: "Ben Parkhurst"
date: "January 28, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)

#Attaching packages
library(tidyverse)
library(janitor)
library(here)
library(readxl)
```
Note that number of # changes the section header size
### Read in inverts.xlsx
```{r}
inverts <- read_excel(here("data","inverts.xlsx"))

```

### `pivot_longer()`
We can use a single colon (:) to indicate a series of sequential columns
```{r}
inverts_long <- pivot_longer(data = inverts,
                            cols = '2016':'2018',
                            names_to = "year",
                            values_to = "sp_count")

#What if you want some in series and some not in the series
#use different chunks

```
What if we want to get it back into wide format?
```{r}
inverts_wide <- inverts_long %>% 
  pivot_wider(names_from = common_name,
              values_from = sp_count)
```
Note: inverts wide is not tidy data.

###`janitor::clean_names()`

```{r}
inverts_wide <- inverts_wide %>% 
  clean_names()

```

###`tidyr::unite()` to combine information from columns

In this example we will merge the site code and year into a single column called "site_year"

```{r}
inverts_unite <- inverts_long %>% 
  unite(col = "site_year",
        c(site, year),
        sep = "_")

```
### Activity:

Create a new object called 'inverts_moyr' starting from inverts_long that unites the month and year columns into a single column names "mo_yr" using a slash "?" as teh separator, then try some other separator.

```{r}
inverts_moyr <- inverts_long %>% 
  unite(col = "mo_yr",
        c(month, year),
        sep = "/")
```
Note: you can merge as many columns as you want separated by the same symbol

### `tidr::separate()`

```{r}
inverts_sep <- inverts_unite %>% 
  separate(col = site_year, into = c("my_site", "my_year"))
```
### `stringer` + `mutate()'

Example: using built in `trees` to add 5 to all of the tree heights
```{r}
trees_new <- trees %>% 
  mutate(height_update = Height + 5)

```
### `str_replace()`

```{r}
ca_abbr <- inverts %>% 
  mutate(common_name = 
           str_replace(common_name,
                       pattern = "california",
                       replacement = "CA"))

  
```
Example: In the common_name column from inverts, you want to replace the word "spiny" with the word "awesome"

Store as `aweomse_lobsters`

```{r}
awesome_lobsters <- inverts %>% 
  mutate(common_name = 
           str_replace(common_name,
                       pattern = "spiny",
                       replacement = "awesome"))
```



